A GRASP algorithm for the concrete delivery problem.

### 2023-08-07 Article/article.tex
Save article

In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic solution to solve a new variant of the CDP. Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site. To the best of our knowledge, this paper is the first that deals with these specific constraints in the context of the CDP.

@@

Regarding objectives, most authors have focused on minimizing travel time and delays between consecutive deliveries. However, some authors have been more interested in maximizing customer satisfaction alone. We find these situations in the works of \cite{durbin2008or, kinable2014concrete, kinable2014logic, sulaman2017simulated}. \cite{kinable2014concrete} introduce a general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet. However, the model did not include constraints limiting the time that concrete may reside in a truck. The authors propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity. The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure. One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batch plants, and 20 concrete mixers. They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances. Finally, the heuristic often yields good solutions in less than a second. A detailed analysis of the MIP model presented in \cite{kinable2014concrete} and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}. In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach. \cite{sulaman2017simulated} expand upon the SD-heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH). This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery. The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots. Experimental results indicated that SATH outperforms SD in speed and solution quality. A generalization of the MIP model of \cite{kinable2014concrete} is addressed in \cite{asbach2009analysis}. This model simultaneously minimizes the total sum of travel costs and the penalty costs for customers with unfulfilled demand. A customer can request that all concrete deliveries come from the same plant or a subset of plants and that a delivery truck belongs to a subset of the vehicle fleet. The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables. \cite{tzanetos2023systematic} provide an overview of the various methods used in the literature to address the CDP and categorizes the problem formulations based on the different concepts used in the literature. They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.

@@

Let $O_i$ be the set of all orders requested by customer $i$. Each element of $O_i$ must be fully delivered before moving on to another order. Exactly one order $o \in O_i$ must have its first delivery start at $a_i$, while the others can start at most $\gamma^2$ time after $o$ is completed. %Once a plant is selected to produce a customer's first delivery, it must be the supplier of all subsequent deliveries from that customer. 
To avoid cold joint problems with the concrete, subsequent deliveries of the same order must be made in close succession. We define a maximum time delay $\gamma^1$ after which no more deliveries are allowed. The customer unloading rate and the quantity to be unloaded give the time required to unload a truckload. Let $\mathcal{C}$ be the set of construction sites (customers) with a planned delivery for the day, and $\mathcal{O}=\{{O_i, i \in \mathcal{C}}\}$ be the set of all requested orders for all customers.

@@

Constraints (\ref{mod:c8}) - - (\ref{mod:c10}) ensure that the first service of any customer $i$ must start at the due time $a_i$. This first service may be performed at the first delivery node of any order $\in O_i$, and one order must be completed before another is started. It also enforces the precedence constraints between the last delivery of an order and the first delivery of the following order.
Constraints (\ref{mod:c11}) - - (\ref{mod:c15}) find the delivery sequence of all orders $\in \mathcal{O}_i$. Constraints (\ref{mod:c16}) - - (\ref{mod:c19}) prohibit two trucks serving consecutive deliveries of the same order from unloading at the same time and enforce a maximum time delay between them. Similarly, constraints (\ref{mod:c20}) - - (\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant. Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities, and (\ref{mod:c240}) bound a driver load.  Constraints (\ref{mod:c25}) require that a driver can only visit a loading/delivery node once. Constraints (\ref{mod:c26}) are degree constraints. Constraints (\ref{mod:c29}) - - (\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service. Finally, constraints (\ref{mod:c31}) - - (\ref{mod:cend}) define the nature and bounds of the variables.

@@

\begin{figure}[H]
    % \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
    \vspace*{-10mm}

@@

\begin{figure}[H]
    \centering

@@

@@

@@

\cite{kinable2014concrete} provided a benchmark dataset for their variant of the $CDP$. Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we can still use this dataset to analyze and compare the performance of our algorithm under different conditions. Table \ref*{tab:cdp_instances} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.  For each method, the table shows the average execution time in milliseconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.  Overall, the GRASP algorithm shows promise in providing near-optimal solutions in a relatively short time for this benchmark of the CDP. It outperformed all methods except CP, achieving an average distance of 5.7\% (13.8\%) with a runtime of 13,186 (574,983) milliseconds for Set A (Set B). The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time. It should also be noted that the exact MIP, CP, and Hybrid are initialized by the results of the SD-heuristic.

@@

>>>>

In this paper, we propose a mathematical model and a Greedy Randomized Search Procedure (GRASP) heuristic to solve a new variant of the CDP. Our model takes into consideration the working shifts of the drivers and the scheduling of multiple orders within the same time window at a construction site. To the best of our knowledge, this paper is the first that deals with these specific constraints in the context of the CDP.

@@

Regarding objectives, most authors have focused on minimizing travel time and delays between consecutive deliveries. However, some authors have been more interested in maximizing customer satisfaction alone. We find these situations in the works of \cite{durbin2008or, kinable2014concrete, kinable2014logic, sulaman2017simulated}. \cite{kinable2014concrete} introduce a general MIP and constraint programming (CP) models of the CDP reflecting the main constraints commonly found in all CDP works: time lag and no overlapping between consecutive deliveries, covering of all customers' demands, delivery time window, and heterogeneous fleet. However, the model did not include constraints limiting the time that concrete may reside in a truck. The authors propose a constructive heuristic that schedules the visits to the customers one by one according to the start time of the visit and the truck capacity. The procedure is invoked multiple times for different permutations of the customer's order which is determined using the steepest descent (SD) local search procedure. One of the paper's main contributions is the creation of the first public test instances for the CDP with up to 50 customers, four batch plants, and 20 concrete mixers. They found the CP model to be highly effective in finding high-quality solutions in a relatively short time or improving existing schedules, while the MIP model can be used to compute bounds, as it seems ineffective in solving large problem instances. Finally, the heuristic often yields good solutions in less than a second. A detailed analysis of the MIP model presented in \cite{kinable2014concrete} and of two more compact models can be found in the thesis of \cite{hernandez_lopez_study_2020}. In \cite{kinable2014logic}, we found an attempt to solve the previous problem with a logic-based Benders' approach. %\cite{sulaman2017simulated} expand upon the SD-heuristic proposed in \cite{kinable2014concrete}, proposing a simulated annealing (SA) combined with a time-slot Heuristic (SATH). This method looks for a slot between existing visits of a truck to schedule a new delivery instead of assigning it to the time slot strictly after the truck's latest assigned delivery. The goal is to reduce the large time gaps that can be present in a schedule created with SD due to ignoring the intermediate available time slots. Experimental results indicated that SATH outperforms SD in speed and solution quality. 
A generalization of the MIP model of \cite{kinable2014concrete} is addressed in \cite{asbach2009analysis}. This model simultaneously minimizes the total sum of travel costs and the penalty costs for customers with unfulfilled demand. A customer can request that all concrete deliveries come from the same plant or a subset of plants and that a delivery truck belongs to a subset of the vehicle fleet. The MIP model is used in a local search scheme as a black-box solver to reoptimize an incumbent solution in which a neighborhood operator has unfixed some variables. \cite{tzanetos2023systematic} provide an overview of the various methods used in the literature to address the CDP and categorizes the problem formulations based on the different concepts used in the literature. They also discussed the consistency between industry needs and existing constraints and provided insights into the datasets corresponding to real-world cases, identifying the necessary data for practitioners.

@@

Let $O_i$ be the set of all orders requested by customer $i$. Each element of $O_i$ must be fully delivered before moving on to another order. Exactly one order $o \in O_i$ must have its first delivery start at $a_i$, while the others can start at most $\gamma^2$ time after $o$ is completed. A plant is assigned to an order, and it must be the supplier of all subsequent deliveries of that order. 
To avoid cold joint problems with the concrete, subsequent deliveries of the same order must be made in close succession. We define a maximum time delay $\gamma^1$ after which no more deliveries are allowed. The customer's unloading rate and the quantity to be unloaded give the time required to unload a truckload. Let $\mathcal{C}$ be the set of construction sites (customers) with a planned delivery for the day, and $\mathcal{O}=\{{O_i, i \in \mathcal{C}}\}$ be the set of all requested orders for all customers.

@@

Constraints (\ref{mod:c8}) - - (\ref{mod:c10}) ensure that the first service of any customer $i$ must start at the due time $a_i$. This first service may be performed at the first delivery node of any order $\in O_i$, and one order must be completed before another is started. It also enforces the precedence constraints between the last delivery of an order and the first delivery of the following order.
Constraints (\ref{mod:c11}) - - (\ref{mod:c15}) find the delivery sequence of all orders $\in \mathcal{O}_i$. With constraints (\ref{mod:c16}) - - (\ref{mod:c19}), two trucks cannot unload at the same time for consecutive deliveries of the same order. Additionally, these constraints impose a maximum time delay between the two trucks. Similarly, constraints (\ref{mod:c20}) - - (\ref{mod:c23}) ensure that two trucks cannot be loaded at the same time at a plant. Constraints (\ref{mod:c24}) require that the cumulative load of all concrete mixers serving an order must equal the required quantities, and (\ref{mod:c240}) bound a driver load.  Constraints (\ref{mod:c25}) require that a driver can only visit a loading/delivery node once. Constraints (\ref{mod:c26}) are degree constraints. Constraints (\ref{mod:c29}) - - (\ref{mod:c30}) calculate the difference between a driver's hours of service and the minimum and normal hours of service. Finally, constraints (\ref{mod:c31}) - - (\ref{mod:cend}) define the nature and bounds of the variables.

@@

\begin{figure}[htbp]
    % \caption{Solution of an instance with two plants, two construction sites, four orders, and three drivers.}
    \vspace*{-10mm}

@@

\begin{figure}[htbp]
    \centering

@@

- resultats avec les instances où toute la demande est satisfaite. Montrer la distance parcourue, le temps de première livraison, overtime, underwork, temps d'execution.

@@

\begin{table}
    \resizebox*{\textwidth}{!}{

\begin{tabular}{lrrrrrrr}
    \toprule
    {} &  satisfiedCost &  driverUsed &  travelCost &  firstDeliveryCost &  underWorkCost &  overTimeCost &  runtime \\
    Name     &                &             &             &                    &                &               &          \\
    \midrule
    20201121 &          226.5 &          13 &     1243.55 &               0.51 &           0.00 &          0.00 &     5.26 \\
    20201010 &          267.0 &          13 &      745.10 &               0.00 &           0.00 &          0.00 &     5.32 \\
    20201107 &          333.5 &          18 &      777.25 &               0.00 &         200.16 &          0.00 &     5.19 \\
    20201031 &          375.0 &          15 &     1017.64 &               0.00 &         216.06 &         28.10 &     5.42 \\
    20201114 &          388.0 &          19 &     1743.64 &               0.00 &         149.48 &          0.00 &     5.42 \\
    20201128 &          776.0 &          31 &     2278.18 &               0.00 &           4.19 &        438.14 &     5.16 \\
    20201024 &          937.5 &          35 &     4403.57 &               0.00 &          58.30 &       1005.24 &     5.72 \\
    20201021 &         1160.5 &          73 &     6107.22 &               0.00 &        2535.40 &        185.02 &     5.28 \\
    20201013 &         1565.0 &         102 &     9214.74 &              19.53 &        2955.35 &         95.74 &     5.32 \\
    20201026 &         1839.5 &         116 &    13585.60 &              44.44 &        1826.70 &        227.56 &     5.41 \\
    20201007 &         2512.5 &         127 &    15166.00 &               5.09 &         877.68 &        574.76 &     5.67 \\
    \bottomrule
    \end{tabular}
    }
\end{table}

@@

\cite{kinable2014concrete} provided a benchmark dataset for their variant of the $CDP$. Although their primary objective is to maximize the total load delivered each day, without considering loading operations at a plant and ensuring deliveries within specified time windows, we can still use this dataset to analyze and compare the performance of our algorithm under different conditions. Table \ref*{tab:cdp_instances} shows the performance of the GRASP heuristic along with the other methods (CP, MIP, SD-heuristic, and Hybrid) used in \cite{kinable2014concrete}.  For each method, the table shows the average execution time in milliseconds and the corresponding gap between the obtained solution and the upper bound provided by \cite{kinable2014concrete}.  Overall, the GRASP algorithm shows promise in providing near-optimal solutions in a relatively short time for this benchmark of the CDP. It outperformed all methods except CP, achieving an average distance of 5.7\% (13.8\%) with a runtime of 13,186 (574,983) milliseconds for Set A (Set B). The SD-heuristic appears to be much faster, but GRASP provides a better balance between solution quality and computation time. It should also be noted that the exact MIP, CP, and Hybrid are initialized with the results of the SD-heuristic.

@@

- rouler les instances sans shift des livreurs. comparer la distance parcourue, et demande livrée.
- rouler sans les livreurs externes. regarder la demande servie par jour.
